Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2020-09-25 manuscript/manuscript.Rmd
More Suparna edits

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "June 8, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past two decades, international nonprofit organizations (INGOs) have faced increasing legal scrutiny abroad, which has negatively impacted public funding flows. Consequently, INGOs may feel the need to fundraise from individuals. Which demographic and experiential factors influence the decision of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint survey experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, but that support evaporates if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do face criticism or crackdown abroad. Our findings have practical implications for INGOs that hope to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

In Human Rights Watch's 2016 annual report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory. Over the past few decades, governments across the globe have been cracking down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" in the policy arena [@CarothersBrechenmacher:2014; @CIVICUS:2017].[^NGOdef] As part of this phenomenon, CSOs, particularly NGOs, face increasing obstacles to their work, including legal restrictions and threats to their freedom of association, expression and speech. Legal restrictions-a core part of these efforts-create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct or repress these organizations-what we refer to as legal crackdowns in this paper.

As a result of these legal crackdowns, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While this has negatively impacted funding flows from public organizations like USAID, DFID, and other OECD funding agencies, philanthropy from private donors, in particular, foundations, appears to not be as adversely affected [@McGill:2018]. Given this trend of closing civic space, NGOs need access to multiple sources of funding and individual-level giving can serve as one such source of funds. What individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy towards their operations abroad?

Existing scholarship on the determinants of individual-level philanthropy to domestic causes is robust across a variety of contexts. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Research on the determinants of philanthropy towards international causes—particularly for organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. Understanding why individuals donate to international causes is essential as it can be harder to motivate as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100].

Individual giving is also important to understand because the amount of money donated by private individuals to CSOs working abroad has steadily increased over time. Total aggregate philanthropy in the United States has grown substantially over time, reaching more than \$400 billion in 2018 (@fig:giving-aggregate). In 2018, total private giving to CSOs working specifically in international affairs reached \$22.08 billion. Moreover, organizations working in the area of international affairs are one of the only two types of recipient organizations that have continued experiencing substantial growth in giving [@givingusa:2019].[^intlaffairs] Most of this is driven by individuals making small-scale or micro-donations.[^small donors] In 2017, individuals gave \$286.7 billion, or approximately 70% of total giving to organizations working in international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [@givingusa:2019]. However, we lack a systematic understanding of which individual donors may be further motivated or deterred to donate under an increasingly restricted civic space abroad. What is the optimal mix of individual-level donor attributes for international NGOs working in restricted political environments interested in maximizing donations? What kinds of individual donors are more or less likely to consider an NGO's host country political environment when making the decision to donate?

![Aggregate philanthopic giving in the United States, 1977–2017. Source: Giving USA.](output/figures/giving_aggregate.pdf){#fig:giving-aggregate}

We explore these questions using a simulated market for philanthropic donations, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing research to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to international NGOs, and that an INGO's target issue area and its relationship with host government interact strongly with these characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with friendly relationships with host governments, preferring instead to donate to NGOs facing more restrictions and working on less contentious issues like emergency response and refugee relief. This may indicate that donors with high trust and support for NGOs could be fair-weather friends that drop their support when their preferred organizations meet resistance abroad.

This study contributes to both to scholarship and practice in public policy, nonprofits and charitable-giving. Individual-level philanthropic giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most NGOs working internationally have previously relied on government and foundation funding. But in the era of closing civic space, these groups may need to readjust their fundraising strategies to cater to individuals. Framing is important in this regard—how people process information affects their willingness to donate.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational-level factors, it can help these CSOs frame their appeals.

More broadly, while research in public policy and nonprofit studies has focused on giving to NGOs, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by international nongovernmental organizations (INGOs)—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, thus shedding light on the changing logic of global governance.

Below, we first examine the state of the existing literature on the key drivers of individual charitable giving, as understanding what motivates private donations is crucial to more effective NGO fundraising. We then describe our survey experiment and simulation methods and explore the results. We conclude with a discussion of the implications our findings have on general INGO funding in the age of constrained civil society space.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive that NGOs have a superior ability to achieve intended outcomes. Therefore, in countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. These groups provide services to large parts of the population in the Global South, especially where governments are unable to do so on a consistent basis.

@@

Thousands of INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds (more commonly known as traditional donor aid) channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles especially from a funding perspective. Official aid donors have reduced the amount of aid channeled to countries with legal restrictions @DupuyPrakash:2018]. In addition, official donors have also redirected funds within restrictive countries, decreasing funds for politically sensitive causes in favor of tamer issues that are more compatible with the governments of other countries [@ChaudhryHeiss:2018]. However, it is notable that private funding towards international nonprofits has not been as adversely affected. Many foundations have continued channeling funds to countries with hostile legal environments [@FoundationCenter2018].

The advent of funding restrictions on NGOs globally and a subsequent decline in government and foundation funding has made the role and value of individual philanthropy in sustaining NGOs even clearer.[^phildef] NGOs need access to a diverse array of funds and individual donations to international affairs organizations have steadily increased over the past two decades. Between 2014-2016, giving to organizations working in international affairs increased by 14.1% [@givingusa2017report]. Moreover, at least in the U.S. context, donations from individuals far outstripped foundation and corporate giving. In 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion from corporations [Source from NVSQ]. Given that a majority of research is restricted to elite high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well. In this paper, we therefore look at which kinds of individual donors—beyond high-net worth donors—are most likely to give abroad.[^highnetworth]

@@

A vast literature on philanthropy looks at organizational-level factors that motivate individuals to donate [@BilodeauSlivinski:1997; @TBPrakash:2017]. However, donation decisions are not determined solely by an organization's characteristics. Donors have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. A robust literature in nonprofit studies has looked at individual donor behavior and characteristics for giving to local—rather than international—causes. This literature focuses on the importance of five main attributes: (1) social capital; (2) the role of higher education; (3) levels of religiosity; (4) wealth; and (5) national group attachment.

One major area of inquiry in philanthropic studies examines the role of social capital, particularly associational capital, or "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Research finds that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not) were both more likely to make charitable donations and make bigger donations. Volunteering exposes individuals to the need for donations, making them feel more compelled to donate [@hossainlamb2012].

Higher education can also play an important role in charitable-giving decisions as it fosters prosocial motivations. It also brings people into particular social networks that entail a higher level of solicitations, which is vital in giving to international causes [@BekkersWiepking:2011]. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations.

Higher education also helps determine the kinds of causes that people donate to globally. A number of NGOs facing crackdowns abroad work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, which is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015].

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions.

Finally, recent research looks at how a sense of ethnonationalism and national group attachment affects the decision to donate to domestic and international charities. Looking at UK respondents, @hartrobson2019 find that whereas nationalists have a preference for domestic charities, they are not necessarily averse to donating to international causes, whereas internationalists have a strong preference for donating to causes and groups working abroad. More specifically, immigrant households are more likely to contribute to international aid organizations [@OktenOsili:2007], and that immigrants can find value in donating to both home and host cultures, especially if their family members still reside in the home country [@burgoyne2005].

Overall, individuals with higher levels of income, education, and greater religious proclivities are not only more likely to give internationally, but also give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. While we expect these determinants to be reflected in our survey experiment, we do not lay out specific hypotheses. Because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Our innovative experimental approach, however, provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

We use a conjoint survey experiment to simultaneously study the effects of organizational characteristics and individual attributes on individuals' preferences to donate to a nonprofit organization. Conjoint analysis methods are well suited for the questions we explore, as they allow us to elicit donor preferences without forcing respondents to explicitly state their exact preferences. Conjoint analysis is commonly used in marketing research to reveal consumer preferences for different mixes of products. It has also been increasingly used in political science and public policy to assess citizen perceptions of political communication [@Helfer:2016; @Mummolo:2016; @KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy.

After collecting baseline information on respondent demographics and attitudes toward charity and voluntarism, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.^[See the appendix for a copy of the survey.] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. These organizations are commonly used in surveys on TODO TODO, in addition to other organizations such as MSF, World Vision (LOOK UP SAVED PAPERS - I THINK NVSQ). We then displayed various combinations of features, including one of four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), one of two organizational practices (financial transparency and third-party accountability audits), one of three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and one of three relationships with host governments (friendly, criticized by the government, and under government crackdown). We presented twelve sets of four hypothetical combinations of organizational features, and because our partial fractional factorial design resulted in 288 (4 × 4 × 2 × 3 × 3) possible combinations of organizations and features, no single respondent was offered every combination.

@@

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and host country politics when deciding to donate. Disentangling the effect of every individual level-characteristics measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates at simultaneously

We use market segmentation methods to construct simulated personas that interact with simulated organizations in a market for charitable contributions. To simplify the many possible combinations of individual characteristics, we generate 32 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes).

@@

Relying on simulation methods offers useful analytic benefits. Rather than interpret individual marginal effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how individual characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, most do not. This is acceptable in simulations, and each combination of persona characteristics is plausible in the real world. Our analysis uses the characteristics of our sample to model the effects of various salient attributes while holding others constant, and thus allows us to explore multiple attributes at once.

Also importantly, our findings are not causal. Rather, they are the descriptive results of simulated data from a nationally representative sample of people who regularly donate to charity. The insights from this simulation, however, still reveal interesting patterns of donor preferences when faced with organizations focusing on a variety of issues, funding sources, and legal restrictions abroad.

@@

Our market simulation reveals which combination of organizational features each persona is most likely to express a preference for donation to. We examine the results of the simulation in three stages, looking at how personas with different (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change their donation preferences across different organizational issue areas, relationships with host governments, and funding sources.

@@

We begin by looking at differences in donation preferences across a range of basic demographic characteristics. The three panels of @fig:demographics shows the average predicted market share for different organizational issue areas, relationships with governments, and funding sources. Though we varied gender across the different personas, gender has no significant effect on any of the simulated outcomes—males and females follow nearly identical patterns across all organizational and individual characteristics. As such, in @fig:demographics, we only vary income. Personas with high income earn \$100,000 a year, while those with less income earn \$50,000 annually.

@@

*(Note: We also plan on examining religiosity and education in the very near future)*

@@

We also combine these dimensions with an additional set of donor characteristics that vary a persona's level of social trust. Personas with high levels of trust have high trust in political institutions and charities, thinks people should be more charitable, frequently volunteers, donates to charity once a month, has a history of political activism, and is a member of an association. Personas with low levels of trust have the opposite characteristics: they do not trust political institutions or charities, rarely donate or volunteer, and are not members or other associations.

@@

@fig:relationship-social shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences, as the average market shares are roughly identical across the two columns of the figure. Similarly, there are no ideological differences across the rows of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. Personas with high levels of social trust are overwhelmingly most likely to donate to organizations with friendly relationships with their host governments, while those with low levels of trust avoid friendly NGOs almost entirely.

@@

![Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different forms of NGO funding](output/figures/funding-social.pdf){#fig:funding-social}

Finally, @fig:funding-social shows the average share of persona market shares across different forms of NGO funding. There are only insignificant differences in donor preferences across public affairs knowledge and experience, political ideology, and social trust. On average, it appears that the source of NGO funding does not play an important role in a donor's preference for the organization. Other organizational factors—in particular issue area and host government relationship—have a much larger influence in shaping donor preferences.

@@

[^NGOdef]: Broadly speaking, NGOs are any not-for-profit, voluntary group organizations at the local, national, or international level. They can either be domestic i.e. an NGO that operates entirely in a single state, or they can be international, more commonly known as INGOs. The latter are composed of members from two or more countries and are organized to advance their members' international goals and provide services to citizens of other states through routine transactions with states, private actors, and international institutions [@Tarrow:2001, 12].

>>>>

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "September 24, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past two decades, international non-governmental organizations (INGOs) have faced increasing legal scrutiny and repression abroad. As a result, official aid to INGOs in repressive countries has reduced. NGOs thus need access to a diverse array of funds and may need to increase fundraising from private donors. Which demographic and experiential factors influence the decision of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint survey experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, but that support evaporates if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do face criticism or crackdown abroad. Our findings have practical implications for INGOs that aim to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

In its 2016 annual report, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory. Governments across the globe have cracked down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017].[^NGOdef] As part of this phenomenon, NGOs face strong legal and bureaucratic obstacles to their work. Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations.

As a result of these legal crackdowns, foreign aid channeled through NGOs has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While this has negatively impacted funding flows from public organizations like United States Agency for International Development (USAID) and Organisation for Economic Cooperation and Development (OECD) funding agencies, philanthropy from private donors, in particular, foundations, appears to not be as adversely affected [@McGill:2018]. Given this trend of closing civic space, INGOs need to maintain access to multiple sources of funding, including individual donors.

Individual private donations are a critical source of INGO funding. Total private philanthropy in the United States has grown substantially over time, reaching more than \$400 billion in 2018 (@fig:giving-aggregate). In 2018, private giving to NGOs working specifically in international affairs reached \$22.08 billion, and international affairs-focused organizations are one of only two categories of organizations that have continued experiencing substantial growth in giving [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale or micro-donations.[^smalldonors] In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018.  But relying on individual donors means that NGOs must consider individual preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad? What kinds of donors are more or less likely to consider an INGO's host country political environment when making the decision to donate?

![Aggregate philanthropic giving in the United States, 1977–2017. Source: Giving USA.](output/figures/giving_aggregate.pdf){#fig:giving-aggregate}

Existing scholarship on the determinants of individual-level philanthropy to domestic causes is robust across a variety of contexts. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. And research on the determinants of philanthropy for international causes—particularly towards organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. While previous work has looked at how organizational attributes of nonprofits affect donors' willingness to donate internationally [@todo:CITEOURPAPERTODIFFERENTIATE), we have yet to uncover demographic traits and experiences that motivate or deter individuals to donate during shrinking civic space.

We explore this question using a simulated market for philanthropic donations, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing research to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to international NGOs, and that an INGO's target issue area and its relationship with host government interact strongly with these characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with friendly relationships with host governments, preferring instead to donate to NGOs facing more restrictions and working on less contentious issues like emergency response and refugee relief. This may indicate that donors with high trust and support for NGOs could be fair-weather friends that drop their support when their preferred organizations meet resistance abroad.

This study contributes to both to scholarship and practice in public policy, nonprofits, and philanthropy. Individual giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most NGOs working internationally have previously relied on government and foundation funding. But with shrinking civic space, these groups may need to readjust their fundraising strategies towards individuals. Framing is important in this regard—how people process information affects their willingness to donate.[^framing] However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help NGOs frame their appeals.

More broadly, while research in public policy examines giving to NGOs, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by INGOs—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance. Thus, it can shed light on the changing logic of global governance.

Below, we first examine the state of the existing literature on the key drivers of individual giving. We then describe our survey experiment and simulation methods and explore the results. We conclude with a discussion of the implications our findings have on general INGO funding in the age of constrained civil society space.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013].

@@

Thousands of INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles from a funding perspective. Official donors have reduced the amount of aid channeled to countries with legal restrictions [@DupuyPrakash:2018] and have redirected funds within restrictive countries, decreasing funds for politically sensitive causes in favor of tamer issues that are more compatible with the governments of those countries [@ChaudhryHeiss:2018]. However, private funding towards international nonprofits has not been as adversely affected. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018].

The advent of funding restrictions on NGOs globally and a subsequent decline in government and foundation funding has made the role and value of individual philanthropy in sustaining NGOs even clearer.[^phildef] NGOs need access to a diverse array of funds and individual donations to international affairs organizations have steadily increased over the past two decades. Moreover, in the United States donations from individuals have far outstripped foundation and corporate giving. In 2018, individual donations amounted to \$292 billion, compared to \$76 billion from foundations and \$20 billion from corporations [@NPT]. Given that a majority of research is restricted to elite high net-worth donors[^highnetworth] [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well.

@@

A vast literature on philanthropy looks at organizational-level factors that motivate individuals to donate [@TBPrakash:2017]. However, donation decisions are not determined solely by a recipient organization's characteristics. Donors have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. Substantial research has looked at the determinants of donating to local—rather than international—causes, including four main attributes: social capital, the role of higher education, levels of religiosity, wealth, and national group attachment.

Social capital, particularly associational capital, or "the networks of formal and informal relationships to which people are associated," is often considered an important indicator of giving [@havens2007, 240]. Volunteering exposes individuals to the need for donations, making them feel more compelled to donate [@hossainlamb2012]. Those who participate in a variety of associations are both more likely to donate as well as make larger donations [@hossainlamb2017].

Higher education also influences charitable-giving decisions as it fosters prosocial motivations. Education brings people into social networks that entail a higher level of solicitations, which is vital in giving to international causes [@BekkersWiepking:2011]. Education also helps determine the kinds of causes that people donate to globally. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations.

A number of NGOs facing crackdowns abroad also work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals. This is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015].

Religiosity is also a significant predictor of the likelihood and level of giving. People who regularly attend religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions.

Finally, recent research looks at how a sense of ethnonationalism and national group attachment affects the decision to donate to domestic and international charities. Looking at UK respondents, @hartrobson2019 find that whereas nationalists have a preference for domestic charities, they are not necessarily averse to donating to international causes, whereas internationalists have a strong preference for donating to causes and groups working abroad.

Which of these characteristics matter when deciding to donate to an INGO, particularly when an organization faces legal crackdown abroad? Overall, individuals with higher levels of income, education, and greater religious proclivities are not only more likely to give internationally, but also give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. Additionally, we posit that other related characteristics such as civic and associational engagement and social trust also influence the decision to donate.

We explore this question using a conjoint survey experiment conducted with a nationally representative sample in the U.S. Importantly, because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Accordingly, we do not test specific hypotheses or theories. However, we do talk about causal effects given the that we conduct an experiment that varies organizational characteristics in order to elicit opinions from respondents (i.e. specific attributes and features cause different aggregate effects). Instead, we structure our analysis around broader themes from the existing literature and investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change their donation preferences across different organizational issue areas, relationships with host governments, and funding sources. Our innovative experimental approach provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

We use a conjoint survey experiment to simultaneously study the effects of organizational characteristics and individual attributes on individuals' preferences to donate to a nonprofit organization. Conjoint analysis allow us to elicit donor preferences without forcing respondents to explicitly state their exact preferences. This kind of analysis is commonly used in marketing research to reveal consumer preferences for different mixes of products, and has become increasingly popular in political science and public policy [@KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy.

In our experiment, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.^[See the appendix for a copy of the survey.] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. Because of their strong name recognition, these organizations (and others like World Vision, the World Wildlife Fund, and Médecins Sans Frontières) are commonly used in experimental surveys on philanthropy [@FaulknerTruongRomaniuk:2015].

```{r child = "output/tables/tbl-organization-attributes-full.md"}
```

We then displayed four hypothetical organizations based on random combinations of features , including one of four issue areas, one of two organizational practices, one of three funding sources, and one of three relationships with host governments (see @tbl:organization-attributes-full; see also Q3.1 in the survey text in the appendix). Respondents saw twelve sets of four hypothetical combinations of organizational features, and because our partial fractional factorial design resulted in 288 (4 × 4 × 2 × 3 × 3) possible combinations of organizations and features, no single respondent was offered every combination. Not every combination necessarily reflected reality—for instance, a respondent might have been asked to compare a fictional Amesty International that focuses on environmental issues and that is funded by private donations with a Greenpeace that focuses on human rights and that is funded by government grants. It is not important that these features align with true organizational characteristics, since the goal of these repeated hypothetical questions is to identify which attributes are the most salient for donors.

@@

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and relationships with host governments when deciding to donate. Disentangling the effect of every individual level-characteristics measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates at simultaneously. To simplify the possible combinations of individual characteristics, we generate 32 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes).

@@

Relying on simulation methods offers useful analytic benefits. Rather than interpret individual marginal effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how individual characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, others may not. This is acceptable in simulations as each combination of persona characteristics is plausible in the real world.

@@

These simulated findings are more abstract than the results from an ordinary survey sample. But by using simulations, we can discover interesting patterns of donor preferences when faced with organizations focusing on a variety of issues, funding sources, and legal restrictions abroad. Our market simulation reveals which combination of organizational features each persona is most likely to express a preference for donation to. We examine the results of the simulation in three stages, looking at how personas with different (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change their donation preferences across different organizational issue areas, relationships with host governments, and funding sources.

@@

We begin by looking at differences in donation preferences across a range of basic demographic characteristics. The three panels of @fig:demographics shows the average predicted market share for different organizational issue areas, relationships with governments, and funding sources. Though we varied gender across the different personas, gender has no significant effect on any of the simulated outcomes—men and women follow nearly identical patterns across all organizational and individual characteristics. As such, in @fig:demographics, we only vary income. Personas with high income earn \$100,000 a year, while those with less income earn \$50,000 annually.

@@

*(TODO: I'll need ≈150 words to describe education and religiosity results)*

@@

We combine these dimensions with an additional set of donor characteristics that vary a persona's level of social trust. Personas with high levels of trust have high trust in political institutions and charities, think people should be more charitable, frequently volunteer, donate to charity once a month, have a history of political activism, and are members of an association. Personas with low levels of trust have the opposite characteristics: they do not trust political institutions or charities, rarely donate or volunteer, and are not members of associations.

@@

@fig:relationship-social shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences, as the average market shares are roughly identical across the two columns of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. Personas with high levels of social trust are overwhelmingly most likely to donate to organizations with friendly relationships with their host governments, while those with low levels of trust avoid friendly NGOs almost entirely.

@@

@@

[^smalldonors]: Micro-donations refer to donations between $0.25–$10; small donors are individuals who donate up to $50 [@SaxtonWang:2014].

[^NGOdef]: Broadly speaking, NGOs are any local, national, or international not-for-profit, voluntary organization. They can be domestic (i.e. an NGO that operates entirely in a single state), or international (INGOs). INGOs are composed of members from two or more countries and are organized to advance their members' international goals and provide services to citizens of other states through routine transactions with states, private actors, and international institutions [@Tarrow:2001, 12].

[^exceptions]: For exceptions, see @Micklewright:2009 on the UK; @Rajan:2009 on Canada; @wiepking2010 on the Netherlands; @CasaleBaumann:2015 on the U.S.
